---
title: "SerumSourceAnalyzer: Heavy/Light Isotope-Based Serum Source Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SerumSourceAnalyzer: Heavy/Light Isotope-Based Serum Source Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 5.5
)
```

```{r setup, message = FALSE}
library(SerumSourceAnalyzer)
library(ggplot2)
```

# 1. Introduction

`SerumSourceAnalyzer` analyzes SILAC-style Heavy/Light dual-channel labeled EV proteomics data. Using UCEV as the control, it classifies proteins in each target group as **serum-derived**:

- **Paired-channel mode** (Case 1/2): the target group's Light fraction significantly exceeds the UCEV background and passes replicate support, CV, and CI checks -> Source-corrected (high/medium confidence, single-peptide candidates);
- **Control-absent mode** (Case 3/4): UCEV has both channels completely missing but the target shows a Light-dominant signal -> Acquired-only (Level 1/2);
- p0 (the UCEV heavy-labeling rate) is used for **source correction** to estimate the serum fraction, and calibrated serum Light intensities are output.

# 2. One-call Full Analysis (real data)

```{r real-data, eval = FALSE}
res <- run_serum_source_analysis(
  input_file = "F:/WYF/NP_1/MaxQuant_annotated_with_total.xlsx",
  output_dir = "F:/WYF/NP_1/Serum_source_analysis_all_groups_1"
)
# Output directory contains: Analysis_Report.md, plots/*.png, all CSV/XLSX result tables
```

# 3. Step-by-Step Pipeline (mini synthetic demo)

Construct a 3-protein x 2-group x 2-replicate mini dataset matching the MaxQuant annotation table structure.

```{r mini-data}
mini <- tibble::tibble(
  `Protein IDs` = c("P_ALB", "P_RES", "P_ACQ"),
  `Representive ID` = c("P_ALB", "P_RES", "P_ACQ"),
  `Protein names` = c("Albumin", "Residual", "Acquired"),
  `Gene names` = c("ALB", "RES", "ACQ"),
  In_house_db = c("in_house_serum", NA, NA),
  Plasma_Altas_db = c("PA001", NA, "PA003"),
  "Intensity L C1" = c(1000, 1000, NA), "Intensity H C1" = c(500, 500, NA),
  "Intensity L C2" = c(1100, 950, NA), "Intensity H C2" = c(550, 480, NA),
  "Intensity L T1" = c(2000, 1000, 1500), "Intensity H T1" = c(400, 480, NA),
  "Intensity L T2" = c(2100, 980, 1400), "Intensity H T2" = c(420, 500, NA),
  "Unique peptides C1" = c(5, 5, NA), "Unique peptides C2" = c(5, 5, NA),
  "Unique peptides T1" = c(5, 5, 3), "Unique peptides T2" = c(5, 5, 3)
)

si <- tibble::tibble(
  sample_id = c("C1", "C2", "T1", "T2"),
  group = c("CTR", "CTR", "TGT", "TGT"),
  light_column = paste("Intensity L", sample_id),
  heavy_column = paste("Intensity H", sample_id),
  peptide_column = paste("Unique peptides", sample_id)
)
```

```{r pipeline}
prep <- prepare_long_data(mini, si)
bg   <- compute_ucev_background(prep$lh_long, "CTR")
res  <- analyze_one_group(prep$lh_long, prep$annotation, "CTR", "TGT",
                          global_fraction_sd_floor = bg$global_fraction_sd_floor)

res %>%
  dplyr::select(Protein_Row_ID, assigned_case, classification,
                UCEV_heavy_labeling_p0, Light_fraction_excess,
                estimated_serum_fraction)
```

The three proteins are classified respectively as:

- `P_ALB`: target Light fraction exceeds control -> **Source-corrected, high confidence**
- `P_RES`: no difference from background -> Ambiguous / Likely_residual
- `P_ACQ`: control completely absent, target has Light-dominant signal -> **Acquired-only (Level 1)**

```{r calibration}
full <- calibrate_serum_light(res, prep$data_converted, si, "TGT", "CTR")
full[, c("Protein_Row_ID", "Serum_L_Calibrated T1", "Serum_L_Calibrated T2")]
```

# 4. Visualization

```{r plots}
cs <- res %>% dplyr::count(comparison, assigned_case, classification,
                           name = "protein_count")
plot_classification_summary(cs)
plot_excess_distribution(res)
plot_excess_vs_probability(res)
plot_serum_fraction(full)
```

# 5. Output System

`run_serum_source_analysis()` produces the following in the output directory:

1. `Analysis_Report.md` -- Markdown report (methods, parameters, summary tables, embedded figures);
2. `plots/` -- 7 PNG figures (classification summary, excess distribution, volcano-style scatter, serum fraction, database evidence, UCEV background, top-protein heatmap);
3. Per-comparison subdirectories + category-wise CSVs, merged tables, database annotation tables;
4. `Serum_Source_Analysis_All_Groups_Summary.xlsx` summary workbook.

Figures can also be saved individually:

```{r save-plot, eval = FALSE}
ggsave("classification.png", plot_classification_summary(res$case_summary),
       width = 9, height = 6, dpi = 300)
```
