---
title: "Heavy/Light Isotope-Based Serum Source Analysis — Execution Notebook"
author: "SerumSourceAnalyzer package"
date: "`r Sys.Date()`"
params:
  INPUT_FILE:  ~
  OUTPUT_DIR:  ~
  PARAMS:      ~
  SAMPLE_INFO: ~
  LIB_DIR:     ~
output:
  html_notebook:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
    code_folding: show
    fig_width: 9
    fig_height: 5
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
    code_folding: show
    self_contained: true
    fig_width: 9
    fig_height: 5
  md_document:
    toc: true
    variant: gfm
    fig_width: 9
    fig_height: 5
    preserve_yaml: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo        = TRUE,
  warning     = TRUE,
  message     = TRUE,
  error       = FALSE,
  comment     = "",
  cache       = FALSE,
  fig.align   = "center"
)
```

# Setup

This notebook reproduces the full Heavy/Light serum-source analysis pipeline
implemented in the `SerumSourceAnalyzer` R package.  It calls the main
`run_serum_source_analysis()` orchestrator (with `render_notebook=FALSE` to
avoid recursion) and then surfaces every intermediate object: console
messages, input shapes, classification summaries, all ggplot figures and
final output tables.  The resulting HTML file is **self-contained** — open
it directly in any modern browser to review the complete execution log.

```{r load-pkg, message=TRUE}
.libPaths(c(params$LIB_DIR, .libPaths()))
suppressPackageStartupMessages({
  library(SerumSourceAnalyzer)
  library(ggplot2)
  library(dplyr)
})
cat("SerumSourceAnalyzer version :", as.character(packageVersion("SerumSourceAnalyzer")), "\n")
cat("Output directory            :", params$OUTPUT_DIR, "\n")
cat("Input workbook              :", basename(params$INPUT_FILE), "\n")
```

---

# Parameters

The user-visible parameter object from `serum_source_params()` and the
sample-to-column mapping used for this run.

```{r params}
input_file  <- params$INPUT_FILE
output_dir  <- params$OUTPUT_DIR
params_obj  <- if (inherits(params$PARAMS, "serum_source_params")) {
  params$PARAMS
} else {
  do.call(serum_source_params, params$PARAMS)
}
sample_info <- if (is.null(params$SAMPLE_INFO)) build_sample_info() else params$SAMPLE_INFO

print(params_obj)
cat("\nGroups to analyse:", paste(unique(sample_info$group), collapse = ", "), "\n")
```

## Sample ↔ MaxQuant column mapping

```{r sample-map}
knitr::kable(
  as.data.frame(sample_info)[, c("sample_id", "group",
                                 "light_column", "heavy_column",
                                 "peptide_column")],
  row.names = FALSE,
  caption = "Sample mapping (en-dash MaxQuant column names)"
)
```

---

# Step 1 — Run the complete pipeline

All reading, preparation, classification, source correction, output saving,
plot generation and Markdown-report writing happens in this one call.  We
set `render_notebook = FALSE` here because **this** Rmd is the notebook
itself; recursion would otherwise loop forever.

```{r step1-run, fig.show='hide', results='hold'}
start_time <- Sys.time()
pipeline_res <- run_serum_source_analysis(
  input_file      = input_file,
  output_dir      = output_dir,
  sample_info     = sample_info,
  params          = params_obj,
  save_outputs    = TRUE,
  make_plots      = TRUE,
  write_report    = TRUE,
  render_notebook = FALSE
)
elapsed <- round(difftime(Sys.time(), start_time, units = "secs"), 1)
cat("Pipeline finished in", as.numeric(elapsed), "seconds.\n")
```

---

# Step 2 — Input & preparation

```{r step2-shapes}
cat("Annotated proteins (rows)           :",
    formatC(nrow(pipeline_res$annotation), format = "d", big.mark = ","), "\n")
cat("Per-protein / per-replicate rows    :",
    formatC(nrow(pipeline_res$lh_long), format = "d", big.mark = ","), "\n")
cat("UCEV proteins (SD estimate base)    :",
    formatC(length(unique(pipeline_res$ucev_sd$Protein_Row_ID)),
            format = "d", big.mark = ","), "\n")
cat("Global Light-fraction SD floor      :",
    signif(pipeline_res$global_fraction_sd_floor, 4), "\n")
```

## First 6 proteins × first 8 annotation columns

```{r step2-head}
knitr::kable(
  utils::head(as.data.frame(pipeline_res$annotation), 6)[, seq_len(min(8, ncol(pipeline_res$annotation)))],
  row.names = FALSE,
  caption = "Head of protein annotation table"
)
```

## Head of per-replicate Heavy/Light table

```{r step2-head-long}
knitr::kable(
  utils::head(as.data.frame(pipeline_res$lh_long), 8),
  row.names = FALSE,
  caption = "Head of the long-format Heavy/Light table"
)
```

---

# Step 3 — Classification summary

Wide-format counts of each classification bucket per comparison group.

```{r step3-summary-wide}
cs <- as.data.frame(pipeline_res$case_summary)
cs$protein_count <- as.integer(cs$protein_count)
wide <- reshape(
  cs[, c("comparison", "classification", "protein_count")],
  idvar     = "comparison",
  timevar   = "classification",
  direction = "wide"
)
names(wide) <- sub("^protein_count\\.", "", names(wide))
wide <- wide[order(wide$comparison), ]
for (cc in setdiff(names(wide), "comparison")) {
  wide[[cc]][is.na(wide[[cc]])] <- 0L
}
knitr::kable(wide, row.names = FALSE,
             caption = "Classification counts per comparison (wide)")
```

## Primary serum-derived protein counts

```{r step3-primary}
prim_count <- pipeline_res$primary_serum %>%
  dplyr::count(comparison, name = "Primary_serum_proteins")
knitr::kable(as.data.frame(prim_count), row.names = FALSE)
```

## Case breakdown by comparison and assigned case

```{r step3-case-long}
knitr::kable(as.data.frame(pipeline_res$case_summary),
             row.names = FALSE,
             caption = "Full case × classification breakdown")
```

---

# Step 4 — Visualisations

Every ggplot object produced by the pipeline is redrawn below so you can
inspect them directly in the notebook.  PNG copies of all these plots also
live under `plots/` inside the output directory.

```{r step4-classification, fig.height=5.2}
pipeline_res$plots$classification_summary
```

```{r step4-excess-dist, fig.height=6}
pipeline_res$plots$excess_distribution
```

```{r step4-excess-vs-prob, fig.height=5.8}
pipeline_res$plots$excess_vs_probability
```

```{r step4-serum-frac, fig.height=5.2}
pipeline_res$plots$serum_fraction
```

```{r step4-database, fig.height=5.0}
pipeline_res$plots$database_evidence
```

```{r step4-ucev-bg, fig.height=4.8}
pipeline_res$plots$ucev_background
```

```{r step4-heatmap, fig.height=10}
pipeline_res$plots$excess_heatmap
```

---

# Step 5 — Database evidence for primary serum proteins

```{r step5-db}
knitr::kable(
  as.data.frame(pipeline_res$database_summary),
  row.names = FALSE,
  caption = "Protein counts split by comparison × classification × database evidence"
)
```

---

# Step 6 — Top 10 serum candidates per group

```{r step6-top10}
top10 <- pipeline_res$primary_serum %>%
  dplyr::group_by(comparison) %>%
  dplyr::arrange(dplyr::desc(Light_fraction_excess), .by_group = TRUE) %>%
  dplyr::slice_head(n = 10) %>%
  dplyr::ungroup() %>%
  dplyr::transmute(
    Group           = comparison,
    Gene            = Gene_names,
    Protein         = substr(Protein_IDs, 1, 40),
    Classification  = classification,
    `Light excess`  = round(Light_fraction_excess, 3),
    `Serum fraction`= round(estimated_serum_fraction, 3),
    `CV (serum fr.)`= round(serum_fraction_CV, 3),
    `Database`      = Database_evidence
  )
knitr::kable(as.data.frame(top10), row.names = FALSE,
             caption = "Top 10 serum-derived candidates per comparison group, ranked by Light-fraction excess")
```

---

# Step 7 — Outputs written to disk

The following files were written inside the output directory.  All paths are
relative to `params$OUTPUT_DIR`.

```{r step7-files}
files <- list.files(output_dir, recursive = FALSE, all.files = FALSE,
                    include.dirs = FALSE)
dirs  <- list.dirs(output_dir, recursive = FALSE, full.names = FALSE)
info  <- data.frame(
  Name     = c(files, dirs),
  Type     = c(rep("file",     length(files)),
               rep("directory", length(dirs))),
  `Size KB`= c(round(file.size(file.path(output_dir, files)) / 1024, 1),
               rep(NA_real_, length(dirs))),
  check.names = FALSE,
  stringsAsFactors = FALSE
)
info <- info[order(info$Type, info$Name), ]
knitr::kable(info, row.names = FALSE,
             caption = "Top-level files produced by the pipeline run")
```

## Location of the Markdown summary report

The plain-Markdown summary report (short version, no code log) is at:

```{r step7-report-ref, results='asis'}
rep_md <- file.path(output_dir, "Analysis_Report.md")
cat("* **Markdown summary** : `", rep_md, "` (",
    ifelse(file.exists(rep_md), paste(round(file.size(rep_md) / 1024, 1), "KB"), "missing"),
    ")\n", sep = "")
```

---

# R session information

For reproducibility — R version, attached packages, OS details.

```{r session-info}
utils::sessionInfo()
```

Total wall-clock time for this notebook run: **`r as.numeric(elapsed)` seconds**.
